Damage localization based on vibrations

### 2017-04-24 rapport/paper.tex
MAJ paper

The purpose of this work is the analysis of a particular data-driven damage localization method,
 namely the FRF interpolation method \cite{dilena2015damage}. Our contribution is the development of a simulator to obtain acceleration measurement of different structures, a proposition of a variant of the FRF interpolation method, and a statistical analysis of the two methods.

@@

Data driven methods are inverse methods that use models based on response data recorded on the structure instead of physical models
\cite{limongelli2016towards}.
Here, a FE model isn't necessary, only recorded data are analyzed.
 In these methods, the damage parameter $D$ is the variation of a damage $d$ between the inspection $I$ and the reference $R$: %Damage-sensitive features are extracted from data and their changes used to identify damage in the structures so the damage parameter D in these method is the variation of the damage feature f between the inspection I and the reference R configurations.
% référence TOWARDS EXTRACTION OF VIBRATION-BASED DAMAGE INDICATORS
% nom pdf : fact sheet template_vibration based_final.pdf

@@

The FRF is evaluated at $s = i \omega$, with $\omega$ the pulsation.
In practice, we want to calculate the FRF of the system, for each sensor localization.
The FRF is calculated for each sensor location and each frequency $\omega_j = \frac{2\pi}{\Delta t} \frac{j}{N}$.
Here, we calculate an estimation of the FRF, called $H_1$ estimator \cite{cauberghe2004applied}. We suppose we have a vector of the forces $\{ f_k\}_{k=1}^N$ and a vector of the accelerations $\{ y_k\}_{k=1}^N$ for each sensor (the two vectors are size $N$). We need to be in frequency domain, so we calculate the discrete Fourier transform:
\begin{equation}
\begin{array}{ll}
F_j = \sum\limits_{k=1}^N f_k e^{-2i\pi kj/N} \ \ \ \ \ j \in \{1 ... N\} \\
Y_j = \sum\limits_{k=1}^N y_k e^{-2i\pi kj/N} \ \ \ \ \ j \in \{1 ... N\}
\end{array}
\end{equation}

@@

Figure \ref{theo} shows this loss of spatial regularity on the theoretical FRF, for a damage at the position 7.
We can see there is a loss of regularity at the position 7 for the damaged state. We can also see there is an irregularity at the position 2, but here the two states have the same irregularity, so it's not a loss of regularity compared to the reference state.

@@

For each sensor, we calculate the Frequency Response Function (FRF) from the acceleration's measures, at certain frequencies
(which are $\frac{k}{N}, k \in \{0 ... N-1\}$ with $N$ the size of the input and output vectors).

@@

Then, we calculate for each sensor $z_l$ the cubic polynomial spline interpolation of the transfer functions $H_R(z_k,j)$ measured at all
the other instrumented locations $\{z_k\}, k\in \{1 ... n\}, k \neq l$: we denote it $H_S(z_l,j)$.
The cubic spline interpolation is a piecewise continuous curve, passing through each of the values.
There is a separate cubic polynomial interpolation for each interval, each with its own coefficients.
This is for this local aspect that we use this interpolation (other interpolations can also be used), since we want to detect local irregularities.
Cubic polynomial spline interpolations produce an interpolated function that is continuous through to the second derivative.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.4\textwidth]{images/interpolation.png}
  \caption{Interpolation error}
  \label{interpolation_error}
\end{figure}

@@

For a decision that an interpolation error belongs to the healthy or damaged state, once have to determine a threshold $E_T$, such as $ \Delta E(z_l)$ must be taller than $E_T$ to have a high probability of identification of a damage at location $z_l$. Indeed, as we can see in the Figure \ref{proba}, the curves which represent the fact to have a damage or not have a share area: if the interpolation error $ \Delta E(z_l)$ is in this area, one can commit two errors:
\begin{itemize}
\item if it is in the red area (FP), once can detect a damage since there isn't one (false alarm).
\item if it is in the green area (FN), once can not detect a damage since there is one (missing alarm).
\end{itemize}

@@

In our experiences, we measure the number of true/false positives and negatives. The Figure \ref{TFPN} present the notations used for the number of true/false positives and negatives.

To evaluate the two methods, we use the specificity predictive factors, which are
$\frac{TP}{TP + FP}$ and $\frac{TN}{TN + FN}$. The first correspond to the probability that there is a damage when we detect one, and the second correspond to the probability there isn't any damage when we detect nothing. A good method has the two factors near to 1.

@@

Now, we have the sampling period:
\begin{equation}
\Delta t = 2 \times max_{j \in \{1 ... n\} } f_j
\end{equation}

@@

From now on, the majority of the figures will show four curves:
\begin{itemize}
\item in dark blue: results of the FRF interpolation method with a threshold at 99\% of the reference errors
\item in green: results of the FRF interpolation method with a threshold at 95\% of the reference distribution
\item in red: results of the FRF difference method with a threshold at 99\% of the reference errors
\item in blue : results of the FRF difference method with a threshold at 95\% of the reference distribution
\end{itemize}

To distinguish the two methods with the two thresholds, we denote them with the color initials: the first one DB, the second one G, the third one R and the last one B.

@@

Figure \ref{detect} presents the percent of detection of a damage between sensors 6 and 7 against the damage extent (a single damage is always simulated on sensor 7). We can see here that below 5\% of damage, none of the methods detect the damage more than 50\% of the time, but we can see that B is the best: at 10\% of damage, it reach 90\% of detection, while the others are below 30\%. However, at 15\% of damage, it's DB and G which are the best, with 100\% of detection. B reaches 100\% of detection at 20\% of damage, and R reaches it at 30\% of damage. Here, R seems to be the worst.

@@

Figure \ref{fp} presents the percent of false positives against the damage percent. Here, R is the best: around 2\% of false positives regardless of the percent of damage, while the others are between 5\% and 25\%.

@@

However, we can see on Figure \ref{detect_noise} that the rate of true positives is impacted by the noise: the G and BD methods reach 100\% of detection at 20\% of damage and the R and BL methods reach it at 40\% of damage.

@@

For the false positives rate, we can notice that the R method is always the best, but the BL method is better than without noise.

@@

Figure \ref{pred} shows the predictive positive and negative factors. For the positive one, we can see that the R method is the best. For the negative one, we can see that all the method reach 1 at 30\% of damage, and are a factor higher than 0.9 for small damages.

@@

We can see that the predictive negative factor isn't bad, but the positive one can be better. To improve it, we must have less false positives. We try to remove them with the next idea: the difference between the threshold and the computed values is larger on the damage than on other sensors. As the damage is localized on one spring, so between two sensors, we apply the next algorithm to determine where is the damage:
\begin{itemize}
\item We compute the error values
\item We make the difference with the threshold
\item We keep only the two highest values, and we detect a damage only if they are positives
\end{itemize}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.4\textwidth]{images/detect_max_10percent.png}
  \caption{}
  \label{detect_max}
\end{figure}

@@

Figure \ref{pred_max} shows the predictive factors with this method. For the positive one, we can see that all the method reach 1 at 30\% of damage. That is what you wanted to obtain with the elimination of the false positives. For the negative one, we can see that this factor is similar at the factor for the initial methods.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.4\textwidth]{images/poutre_curve2.png}
  \caption{}
  \label{beam_curve}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.4\textwidth]{images/poutre_error.png}
  \caption{}
  \label{beam_error}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.4\textwidth]{images/quantification.png}
  \caption{}
  \label{quant}
\end{figure}

@@

In this paper, we presented the structure health monitoring and a state of the art for this. Then we detailed one of the methods, the FRF interpolation method, and some physical models. Last, we presented our simulation and our results with the implementation of the FRF interpolation method, and analyzed them.

>>>>

The purpose of this work is the analysis of a particular data-driven damage localization method,
 namely the FRF interpolation method \cite{dilena2015damage}.
%Our contribution is the development of a simulator to obtain acceleration measurement of different structures, a proposition of a variant of the FRF interpolation method, and a statistical analysis of the two methods.
Our contribution is a statistical evaluation of the method. This led to a proposition of a variant of the FRF interpolation method, and both methods were analyzed in extensive simulations.

@@

Data driven methods are inverse methods that use models based on response data recorded on the structure instead of physical models
\cite{limongelli2016towards}.
Here, a FE model is not necessary, only recorded data are analyzed.
 In these methods, the damage parameter $D$ is the variation of a damage $d$ between the inspection $I$ and the reference $R$: %Damage-sensitive features are extracted from data and their changes used to identify damage in the structures so the damage parameter D in these method is the variation of the damage feature f between the inspection I and the reference R configurations.
% référence TOWARDS EXTRACTION OF VIBRATION-BASED DAMAGE INDICATORS
% nom pdf : fact sheet template_vibration based_final.pdf

@@

The FRF is evaluated at $s = i \omega$, with $\omega$ the pulsation.
In practice, we want to calculate the FRF of the system, for each sensor location.
The FRF is calculated for each sensor location and each frequency $\omega_j = \frac{2\pi}{\Delta t} \frac{j}{N}$.
Here, we calculate an estimation of the FRF, called $H_1$ estimator \cite{cauberghe2004applied}. We suppose we have a vector of the forces $\{ f_k\}_{k=1}^N$ and a vector of the accelerations $\{ y_k\}_{k=1}^N$ for each sensor (the two vectors are size $N$). We need to be in frequency domain, so we calculate the discrete Fourier transform:
\begin{equation}
\begin{array}{ll}
F_j = \sum\limits_{k=1}^N f_k e^{-2i\pi kj/N} \ \ \ \ \ j \in \{1 ... N\} \\
Y_j = \sum\limits_{k=1}^N y_k e^{-2i\pi kj/N} \ \ \ \ \ j \in \{1 ... N\}
\end{array}
\end{equation}

@@

Figure \ref{theo} shows this loss of spatial regularity on the theoretical FRF, for a damage at the position 7.
We can see there is a loss of regularity at the position 7 for the damaged state. We can also see there is an irregularity at the position 2, but here the two states have the same irregularity, so it is not a loss of regularity compared to the reference state.

@@

For each sensor, we calculate the Frequency Response Function (FRF) from the acceleration's measures, at certain frequencies
(which are $\frac{j}{N}, j \in \{0 ... N-1\}$ with $N$ the length of the input and output vectors).

@@

Then, we calculate for each sensor $z_l$ the cubic polynomial spline interpolation from the transfer functions $H_R(z_k,j)$ measured at all
the other instrumented locations $\{z_k\}, k\in \{1 ... n\}, k \neq l$: we denote it $H_S(z_l,j)$.
The cubic spline interpolation is a piecewise continuous curve, passing through each of the values.
There is a separate cubic polynomial interpolation for each interval, each with its own coefficients.
This is for this local aspect that we use this interpolation (other interpolations can also be used), since we want to detect local irregularities.
Cubic polynomial spline interpolations produce an interpolated function that is continuous through to the second derivative.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.4\textwidth]{images/interpolation2.png}
  \caption{Interpolation error E}
  \label{interpolation_error}
\end{figure}

@@

For a decision that an interpolation error belongs to the healthy or damaged state, a threshold $E_T$ has to be determined, such as $ \Delta E(z_l)$ must be taller than $E_T$ to have a high probability of identification of a damage at location $z_l$. Indeed, as we can see in the Figure \ref{proba}, the curves which represent the fact to have a damage or not may share an area: if the interpolation error $ \Delta E(z_l)$ is in this area, one can commit two errors:
\begin{itemize}
\item if it is in the red area (FN), a damage is present but not detected (false alarm).
\item if it is in the green area (FP), a damage is detected while there is actually no damage (missing alarm).
\end{itemize}

@@

In our experiments, we measure the number of true/false positives and negatives. The Figure \ref{TFPN} present the notations used for the number of true/false positives and negatives.

To evaluate the two methods, we use the specificity predictive factors, which are
$\frac{TP}{TP + FP}$ and $\frac{TN}{TN + FN}$. The first correspond to the probability that there is a damage when we detect one, and the second correspond to the probability there is not any damage when we detect nothing. A good method has the two factors near to 1.

@@

Now, we have the sampling period $\Delta t$ which is given by:
\begin{equation}
\frac{1}{\Delta t} \geq 2 \times max_{j \in \{1 ... n\} } f_j
\end{equation}

@@

From now on, the majority of the figures will show four curves:
\begin{itemize}
\item in blue (continuous line): results of the FRF interpolation method with a threshold at 99\% of the reference errors
\item in blue (dashes): results of the FRF interpolation method with a threshold at 95\% of the reference distribution
\item in red (continuous line): results of the FRF difference method with a threshold at 99\% of the reference errors
\item in red (dashes): results of the FRF difference method with a threshold at 95\% of the reference distribution
\end{itemize}

To distinguish the two methods with the two thresholds, we denote them as follows: the first one I99, the second one I95, the third one D99 and the last one D95.

@@

Figure \ref{detect} presents the percent of detection of a damage between sensors 6 and 7 against the damage extent (a single damage is always simulated on sensor 7). We can see here that below 5\% of damage, none of the methods detect the damage more than 50\% of the time, but we can see that D95 is the best: at 10\% of damage, it reach 90\% of detection, while the others are below 30\%. However, at 15\% of damage, it is I95 which is the best, with 99\% of detection. I99 and D95 reach 100\% of detection at 20\% of damage, and D99 reaches it at 30\% of damage. Here, D99 seems to be the worst.

@@

Figure \ref{fp} presents the percent of false positives against the damage percent. Here, D99 is the best: around 2\% of false positives regardless of the percent of damage, while the others are between 5\% and 25\%.

@@

However, we can see on Figure \ref{detect_noise} that the rate of true positives is impacted by the noise: the I99 and I95 methods reach 100\% of detection at 20\% of damage and the D99 and D95 methods reach it at 40\% of damage.

@@

For the false positives rate, we can notice that the D99 method is always the best, but the D95 method is better than without noise.

@@

Figure \ref{pred} shows the predictive positive and negative factors. For the positive one, we can see that the D99 method is the best. For the negative one, we can see that all the method reach 1 at 30\% of damage, and are a factor higher than 0.9 for small damages.

@@

We can see that the predictive negative factor is not bad, but the positive one can be better. To improve it, we must have less false positives. We try to remove them with the next idea: the difference between the threshold and the computed values is larger on the damage than on other sensors. As the damage is localized on one spring, so between two sensors, we apply the next algorithm to determine where is the damage:
\begin{itemize}
\item We compute the error values
\item We make the difference with the threshold
\item We keep only the two highest values, and we detect a damage only if they are positives
\end{itemize}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.4\textwidth]{images/detect_max_30percent.png}
  \caption{Damage detection rate for a damage (30\%) between sensors 6 and 7, with elimination of false positives}
  \label{detect_max}
\end{figure}

@@

Figure \ref{pred_max} shows the predictive factors with this method. For the positive one, we can see that all the method reach 1 at 30\% of damage. That is what you wanted to obtain with the elimination of the false positives. For the negative one, we can see that this factor is similar at the factor for the initial methods. Here, the method with the best results is the I95 method.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.4\textwidth]{images/poutre_curve2.png}
  \caption{FRF for undamaged and damage states of the beam at one frequency, against sensors}
  \label{beam_curve}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.4\textwidth]{images/poutre_error.png}
  \caption{Interpolation error with the beam}
  \label{beam_error}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.4\textwidth]{images/quantification.png}
  \caption{Error (in percent of the threshold)}
  \label{quant}
\end{figure}

@@

In this paper, we presented the structural health monitoring and a state of the art for this. Then we detailed one of the methods, the FRF interpolation method, and the underlying physical models. Last, we presented our simulation and our results with the implementation of the FRF interpolation method, and analyzed them.
